Parameters estimation via HOSVD

### 2025-08-23 Code/R/Research/HOSVD parameters estimate.Rmd
CDAM comparison results

## Графики RMSE
```{r}
pltnms <- c("_no_rates", "_small_eq_rates", "_large_eq_rates", "")
pn <- 4

load(file = paste0("./Research/comp_results/signal", pltnms[pn], "_hossa.RData"))
load(file = paste0("./Research/comp_results/signal", pltnms[pn], "_ssa.RData"))
```

```{r}
rec.df <- rec.rmse |> as.data.frame() |>
  mutate(dims = factor(x = apply(tens.dims, 1, paste, collapse = "x"),
                       levels = apply(tens.dims, 1, paste, collapse = "x"))) |>
  pivot_longer(!dims, names_to = "trunc dims", names_prefix = "V", values_to = "RMSE") |>
  mutate(`trunc dims` = as.factor(paste(trunc_dims[as.integer(`trunc dims`)])))

rec.plot <- rec.df |> ggplot() +
  geom_line(aes(
    x = dims,
    y = RMSE,
    group = `trunc dims`,
    colour = `trunc dims`,
    linetype = `trunc dims`
  ), lwd = 1.2) + 
  geom_abline(
    slope = 0,
    intercept = min.ssa.rmse[1],
    col = "gray10",
    lty = 3
  ) +
  ylim(min.ssa.rmse[1] * 0.985, NA) +
  annotate(
    geom = "text",
    x = 5,
    y = min.ssa.rmse[1] * 0.99,
    label = paste0("SSA min RMSE, L=", min.ssa.rmse[2])
  ) +
  theme(axis.text.x = element_text(angle = 90))

plot.save(
  rec.plot,
  paste0("./img/rec_dim_rmse", pltnms[pn], ".pdf"),
  pdf,
  width = 9,
  height = 6,
  save = save
)
```

@@

```{r}
freq1.df <- freq1.rrmse |> as.data.frame() |>
  mutate(dims = factor(
    x = apply(tens.dims, 1, paste, collapse = "x"),
    levels = apply(tens.dims, 1, paste, collapse = "x")
  )) |>
  pivot_longer(
    !dims,
    names_to = "Estimation\nDimension",
    names_prefix = "V",
    values_to = "RRMSE"
  )

freq1.plot <- freq1.df[-(1:(2 * 3)), ] |> ggplot() +
  geom_line(
    aes(
      x = dims,
      y = RRMSE,
      group = `Estimation\nDimension`,
      colour = `Estimation\nDimension`,
      linetype = `Estimation\nDimension`
    ),
    lwd = 1.2
  ) +
  geom_abline(
    slope = 0,
    intercept = min.ssa.rrmse$freq1[1],
    col = "gray10",
    lty = 3
  ) +
  annotate(
    geom = "text",
    x = 7,
    y = min.ssa.rrmse$freq1[1] * 0.95,
    label = paste0("ESPRIT min RRMSE, L=", min.ssa.rrmse$freq1[2])
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90))

@@

# no rates
# rates.true <- c(0, 0)
# plot.name <- "_no_rates"
# both rates -0.01
# rates.true <- c(-0.01, -0.01)
# plot.name <- "_small_eq_rates"
# both rates -0.02
rates.true <- c(-0.02, -0.02)
plot.name <- "_large_eq_rates"
# rates from paper
# rates.true <- c(-0.01, -0.02)
# plot.name <- ""

@@

## Графики RRMSE оценок параметров каждым из методов
```{r}
pltnms <- c("_no_rates", "_small_eq_rates", "_large_eq_rates", "")
pn <- 4

@@

```{r}
freq1.df <- freq1.mult.rrmse |> as.data.frame() |> mutate(L = Ls) |>
  pivot_longer(!L, names_to = "Method", values_to = "RRMSE") |>
    mutate(Method = ifelse(Method == "HO.MSSA", "HO-M-ESPRIT", "M-ESPRIT"))
freq1.plot <- freq1.df |> ggplot() + geom_line(aes(
  x = L,
  y = RRMSE,
  colour = Method,
  linetype = Method
), lwd = 1.2)

plot.save(
  freq1.plot,
  paste0("./img/freq1_L", pltnms[pn], ".pdf"),
  pdf,
  width = 9,
  height = 6,
  save = save
)
```

```{r}
freq2.df <- freq2.mult.rrmse |> as.data.frame() |> mutate(L = Ls) |>
  pivot_longer(!L, names_to = "Method", values_to = "RRMSE") |>
    mutate(Method = ifelse(Method == "HO.MSSA", "HO-M-ESPRIT", "M-ESPRIT"))
freq2.plot <- freq2.df |> ggplot() + geom_line(aes(
  x = L,
  y = RRMSE,
  colour = Method,
  linetype = Method
), lwd = 1.2)

plot.save(
  freq2.plot,
  paste0("./img/freq2_L", pltnms[pn], ".pdf"),
  pdf,
  width = 9,
  height = 6,
  save = save
)
```

@@

## Графики RMSE
```{r}
pltnms <- c("_no_rates", "_small_eq_rates", "_large_eq_rates", "")
pn <- 4

@@

rec.plot <- rec.df |> ggplot() +
  geom_line(aes(
    x = L,
    y = RMSE,
    group = Method,
    colour = Method,
    linetype = Method
  ), lwd = 1.2)

plot.save(
  rec.plot,
  paste0("./img/rec_L_rmse", pltnms[pn], ".pdf"),
  pdf,
  width = 9,
  height = 6,
  save = save
)
```

>>>>

## Графики RMSE
```{r}
pltnms <- c("_no_rates", "_small_eq_rates", "_large_eq_rates", "")
pn <- 1

load(file = paste0("./Research/comp_results/signal", pltnms[pn], "_hossa.RData"))
load(file = paste0("./Research/comp_results/signal", pltnms[pn], "_ssa.RData"))
min.ssa.rmse <- c(min(rec.ssa.rmse), Ls[which.min(rec.ssa.rmse)])
```

```{r}
rec.df <- rec.rmse |> as.data.frame() |>
  mutate(dims = factor(x = apply(tens.dims, 1, paste, collapse = "x"),
                       levels = apply(tens.dims, 1, paste, collapse = "x"))) |>
  pivot_longer(!dims, names_to = "trunc dims", names_prefix = "V", values_to = "RMSE") |>
  mutate(`trunc dims` = as.factor(paste(trunc_dims[as.integer(`trunc dims`)]))) |> 
  filter(`trunc dims` == "1")

rec.plot <- rec.df |> ggplot() +
  geom_line(aes(
    x = dims,
    y = RMSE,
    group = `trunc dims`,
    # group = `trunc dims`,
    # colour = `trunc dims`,
    # linetype = `trunc dims`
  ), lwd = 1.2) + 
  geom_abline(
    slope = 0,
    intercept = min.ssa.rmse[1],
    col = "gray10",
    lty = 3
  ) +
  ylim(min.ssa.rmse[1] * 0.985, NA) +
  annotate(
    geom = "text",
    x = 5,
    y = min.ssa.rmse[1] * 0.99,
    label = paste0("SSA min RMSE, L=", min.ssa.rmse[2])
  ) + theme_minimal() +
  theme(axis.text.x = element_text(angle = 90))

plot.save(
  rec.plot,
  paste0("./img/rec_dim_rmse", pltnms[pn], ".pdf"),
  pdf,
  width = 11,
  height = 5,
  save = save
)
```

@@

```{r}
freq1.df <- freq1.rrmse |> as.data.frame() |>
  mutate(dims = factor(
    x = apply(tens.dims, 1, paste, collapse = "x"),
    levels = apply(tens.dims, 1, paste, collapse = "x")
  )) |>
  pivot_longer(
    !dims,
    names_to = "Estimation\nDimension",
    names_prefix = "V",
    values_to = "RRMSE"
  ) |> mutate(RMSE = RRMSE * freqs.true[1] / 100)

freq1.plot <- freq1.df[-(1:(2 * 3)), ] |> ggplot() +
  geom_line(
    aes(
      x = dims,
      y = RMSE,
      group = `Estimation\nDimension`,
      colour = `Estimation\nDimension`,
      linetype = `Estimation\nDimension`
    ),
    lwd = 1.2
  ) +
  geom_abline(
    slope = 0,
    intercept = min.ssa.rrmse$freq1[1] * freqs.true[1] / 100,
    col = "gray10",
    lty = 3
  ) +
  annotate(
    geom = "text",
    x = 7,
    y = min.ssa.rrmse$freq1[1] * 0.95 * freqs.true[1] / 100,
    label = paste0("ESPRIT min RMSE, L=", min.ssa.rrmse$freq1[2])
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90))

@@

@@

@@

```{r}
freq1.df <- freq1.mult.rrmse |> as.data.frame() |> mutate(L = Ls) |>
  pivot_longer(!L, names_to = "Method", values_to = "RRMSE") |>
    mutate(Method = ifelse(Method == "HO.MSSA", "HO-M-ESPRIT", "M-ESPRIT"), 
           RMSE = RRMSE * abs(freqs.true[1]) / 100)
freq1.plot <- freq1.df |> ggplot() + geom_line(aes(
  x = L,
  y = RMSE,
  colour = Method,
  linetype = Method
), lwd = 1.2) + theme_minimal()

plot.save(
  freq1.plot,
  paste0("./img/freq1_L", pltnms[pn], ".pdf"),
  pdf,
  width = 11,
  height = 5,
  save = save
)
```

```{r}
freq2.df <- freq2.mult.rrmse |> as.data.frame() |> mutate(L = Ls) |>
  pivot_longer(!L, names_to = "Method", values_to = "RRMSE") |>
    mutate(Method = ifelse(Method == "HO.MSSA", "HO-M-ESPRIT", "M-ESPRIT"),
           RMSE = RRMSE * abs(freqs.true[2]))
freq2.plot <- freq2.df |> ggplot() + geom_line(aes(
  x = L,
  y = RMSE,
  colour = Method,
  linetype = Method
), lwd = 1.2) + theme_minimal()

plot.save(
  freq2.plot,
  paste0("./img/freq2_L", pltnms[pn], ".pdf"),
  pdf,
  width = 11,
  height = 5,
  save = save
)
```

@@

## Графики RMSE
```{r}
pltnms <- c("_no_rates", "_small_eq_rates", "_large_eq_rates", "")
pn <- 1

@@

rec.plot <- rec.df |> ggplot() +
  geom_line(aes(
    x = L,
    y = RMSE,
    group = Method,
    colour = Method,
    linetype = Method
  ), lwd = 1.2) + theme_minimal()

plot.save(
  rec.plot,
  paste0("./img/rec_L_rmse", pltnms[pn], ".pdf"),
  pdf,
  width = 11,
  height = 5,
  save = save
)
```
